### 2024-09-21 Paper/Paper.qmd

---
title: "My title"
subtitle: "My subtitle if needed"
author: Krishna Kumar
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

# Data

-   Can add summary stats of the data for the reader to get a better idea

-   Like just have the average, standard deviation, maybe a QQ plot or smth

# Results

# Discussion

# References